---
title: "Winning Jeopardy"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
jeopardy <- read_csv('jeopardy.csv')
print(head(jeopardy, 5))
print(colnames(jeopardy))
```

```{r}
colnames(jeopardy) <- c('show_number', 'air_date', 'round', 'category', 'value', 'question', 'answer')
print(sapply(jeopardy, typeof))
```

```{r}
jeopardy <- jeopardy %>%
  filter(value != 'None') %>%
  mutate(value = str_replace_all(value, '[$,]', ''),
         value = as.numeric(value))
```

```{r}
jeopardy <- jeopardy %>%
  mutate(
    question = tolower(question),
    question = str_replace_all(question, '[^A-Za-z0-9 ]', ''),
    answer = tolower(answer),
    answer = str_replace_all(answer, '[^A-Za-z0-9 ]', ''),
    category = tolower(category),
    category = str_replace_all(category, '[^A-Za-z0-9 ]', '')
  )
```

```{r}
jeopardy <- jeopardy %>%
  separate(., air_date, into = c('year', 'month', 'day'), sep='-') %>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    day = as.numeric(day)
  )
```

```{r}
n_questions <- nrow(jeopardy)
p_category_expected <- 1/3369
p_not_category_expected <- 3368/3369
p_expected <- c(p_category_expected, p_not_category_expected)
```

```{r}
categories <- pull(jeopardy, category)
n_science_categories <- 0

for (c in categories) {
  if ('science' %in% c) {
    n_science_categories = n_science_categories + 1
  }
}

science_observations = c(n_science_categories, n_questions - n_science_categories)
chisq.test(science_observations, p = p_expected)
```

Given such a small p-value, we should reject the null hypothesis that science does not occur more frequently than other topics.

```{r}
n_history_categories <- 0

for (c in categories) {
  if ('history' %in% c) {
    n_history_categories = n_history_categories + 1
  }
}

history_observations = c(n_history_categories, n_questions - n_history_categories)
chisq.test(history_observations, p = p_expected)
```

Again, the p-value is so small that we would reject the null hypothesis that history questions occur more frequently than others.

```{r}
n_shakespeare_categories <- 0

for (c in categories) {
  if ('shakespeare' %in% c) {
    n_shakespeare_categories = n_shakespeare_categories + 1
  }
}

shakespeare_observations = c(n_shakespeare_categories, n_questions - n_shakespeare_categories)
chisq.test(shakespeare_observations, p = p_expected)
```

Finally, we would again reject the null hypothesis that Shakespeare questions occur more frequently than other questions.

# Investigating the repetition of old questions

```{r}
jeopardy <- jeopardy %>%
  arrange(year, month, day)

questions <- pull(jeopardy, question)
terms_used <- character(0)

for (q in questions) {
  split_sentence <- str_split(q, ' ')[[1]]
  
  for (term in split_sentence) {
    if (!term %in% terms_used & nchar(term) >= 6) {
      terms_used = c(terms_used, term)
    }
  }
}
```

# High Value vs. Low Value Terms

```{r}
values <- pull(jeopardy, value)
value_count = NULL

for (term in terms_used[1:20]) {
  n_high_value = 0
  n_low_value = 0
  
  for (i in 1:length(questions)) {
    split_sentence = str_split(questions[i], ' ')[[1]]
    
    if (term %in% split_sentence & values[i] >= 800) {
      n_high_value = n_high_value + 1
    } else if (term %in% split_sentence & values[i] < 800) {
      n_low_value = n_low_value + 1
    }
  }
  
  test <- chisq.test(c(n_high_value, n_low_value), p = c(2/5, 3/5))
  new_row <- c(term, n_high_value, n_low_value, test$p.value)
  
  value_count <- rbind(value_count, new_row)
}
```

```{r}
tidy_value_count <- as_tibble(value_count)
colnames(tidy_value_count) <- c('term', 'n_high', 'n_low', 'p_value')

head(tidy_value_count, 20)
```

We received a ton of warnings most likely due to where we have values less than 5. These are probably not useful values. Taking a smaller sample of results, we can see that american, minister and created seem to be more associated with high value questions.